---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# maskr

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/maskr)](https://CRAN.R-project.org/package=maskr)
<!-- badges: end -->

The `maskr` package provides a masked array type, which is either a vector, array, matrix, or data frame which has a *mask*.
The mask is a list of indices, one for each axis of the array, which defines the parts of the array to show. 
All other parts then become invisible.

## Installation

You can install the development version of maskr from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("multimeric/maskr")
```

## Basics

You can create a `MaskedArray` using the constructor function, which takes firstly
the array to mask, and secondly a list of mask vectors, one for each dimension of the array.
For a masked vector, the mask is a list of one vector.

Here we define an inner array with 10 elements, but our mask, which is a vector of indices to show, hides `"a"` and `"j"`:

```{r example}
x = maskr::MaskedArray(letters[1:10], masks=list(2:9))
x
```

For most purposes, the array acts the same as `c("b", "c", "d", "e", "f", "g", "h", "i")`:

```{r}
length(x)
```
```{r}
x[[1]]
```

However, we can unmask the array the return to the original form:

```{r}
maskr::unmask(x)
```

## Use Cases

One use case for `maskr` is hiding some kind of bad data, like outliers.

For example, let's start with some data that includes two outliers

```{r}
set.seed(1)
x = c(
  rnorm(10),
  rnorm(2, mean=2)
)
x
```

We could identify these outliers using some kind of test. 
In this case we test that all the values come from from a $N(0, 1)$ distribution:

```{r}
outliers = pnorm(x, lower.tail = F) < 0.05
outliers
```

Then we can use that outlier list as a mask. 
Note that logical vectors are also legal masks!

```{r}
masked = maskr::MaskedArray(x, mask = list(!outliers))
masked
```

We might then want to find the mean of the non-outliers:

```{r}
mean(masked)
```

## Subsetting

Another useful feature of `maskr` is that, when you subset an array, the mask shrinks to hide the excluded elements:

```{r}
x[5:6]
```

What this means, is that you can take a sequence of slices from the array:

```{r}
y = x[-1][-1][-1]
```

We can compute the final subset:
```{r}
maskr::apply_mask(y)
```
But also revert the mask:
```{r}
maskr::unmask(y)
```

## Multidimensional Arrays

We can mask a matrix. If we do, the mask needs to be a list of two vectors:

```{r}
x = maskr::MaskedArray(matrix(rnorm(n=16), nrow=4), list(1:2, 2:3))
x
```
```{r}
maskr::apply_mask(x)
```

We can also mask a `data.frame` in the same way. 
Note that we can use a character vector to select certain columns, as you might expect:

```{r}
x = maskr::MaskedArray(iris, list(10:15, c("Petal.Length", "Sepal.Length")))
maskr::apply_mask(x)
```
